Mathematical Aspects of the Capital Market

### 2017-08-07 phd/Chapterintroduction.tex
Finished the description of the algorithm.

\section{Stochastic Recurrence Equation}
One of the most important dynamical mechanisms that lead to regularly
varying r.v. is stochastic recursion of the following form:
\begin{equation}
  \label{eq:rhjyu}
  X_t = A_t X_{t-1} + B_t
\end{equation}
where $X_t$ is a $d$-dimensional random vector, $A_t$ is a $d\times d$
random matrix and $B_t$ is a $d$-dimensional vector, random or
deterministic. The sequence $\{A_i\}_{i=1,2,\dots}$ and
$\{B_i\}_{i=1,2,\dots}$  are iid and independent of each other.

@@

\section{GARCH models}
Introduced by Bollerslev in 1986, GARCH models have been hugely
popular in modelling volatility of financial time series and have
inspired numerous variants. Its original form is the following the
recurrence equation:
\[
\sigma_t^2 = \omega + \sum_{i=1}^p \alpha_i R_{t-i}^2 +
\sum_{j=1}^q \beta_j \sigma_{t-j}^2
\]
where $\R_t$ is a return series, e.g. stock returns, foreign exchange
rates, interest rates, etc; $\sigma_t^2$ is the variance of the
distribution of $R_t$ conditional on $\{(R_i,
\sigma_i^2)\}_{i=0}^{t-1}$. $\omega, \{\alpha_i\}_{i=1}^p,
\{\beta_i\}_{i=1}^q$ are constant parameters of the model. Clearly,
the GARCH($p,q$) recurrence equation is of the form of
\eqref{eq:rhjyu}. So, with appropriate conditions, including
$\sum_{i=1}^p \alpha_i + \sum_{j=1}^q \beta_j < 1$, $\sigma_t^2$ is
shown to be a positive Harris Markov chain (cf. Bollerslev
\cite{bollerslev:1986} and Buraczewski et al
\cite{buraczewski:damek:mikosch:2016}), whose stationary distribution
has regularly varying tails. The tail index, call it $\xi$, is given by
\[
\lim_{n \to \infty} {1 \over n}\log\E\|A_n \cdots A_1\|^\xi = 0
\]
where $A_1, A_2, \dots, A_n$ are iid matrices whose entries are
functions of $\{\alpha_i\}_{i=1}^p$ and $\{\beta_i\}_{i=1}^q$:
\[
A_i =
\begin{pmatrix}
  \alpha_1 Z_{t-1}^2 + \beta_1 & \beta_2 & \cdots &
  \beta_{q-1} & \beta_q & \alpha_2 & \alpha_3 &
  \cdots & \alpha_{p-1} & \alpha_p\\
  1 & 0 & \cdots & 
  0 & 0 & 0 & 0 & \cdots & 0 & 0 \\
  \vdots & \vdots & \ddots & 
  \vdots & \vdots & \vdots & \vdots &
  \ddots & \vdots & \vdots \\
  0 & 0 & \cdots &
  0 & 0 & 0 & 0 & \cdots & 0 & 0 \\
  0 & 0 & \cdots &
  1 & 0 & 0 & 0 & \cdots & 0 & 0 \\
  Z_{t-1}^2 & 0 & \cdots &
  0 & 0 & 0 & 0 & \cdots & 0 & 0 \\
  0 & 0 & \cdots &
  0 & 0 & 1 & 0 & \cdots & 0 & 0 \\
  \vdots & \vdots & \ddots &
  \vdots & \vdots & \vdots & \vdots &
  \ddots & \vdots & \vdots \\
  0 & 0 & \cdots &
  0 & 0 & 0 & 0 & \cdots & 0 & 0 \\    
  0 & 0 & \cdots &
  0 & 0 & 0 & 0 & \cdots & 1 & 0 \\    
\end{pmatrix}
\begin{pmatrix}
  \sigma_{t-1}^2 \\
  \sigma_{t-2}^2 \\
  \vdots \\
  \sigma_{t-q+1}^2 \\
  \sigma_{t-q}^2 \\
  R_{t-2}^2 \\
  R_{t-3}^2 \\
  \vdots \\
  R_{t-p+1}^2 \\
  R_{t-p}^2
\end{pmatrix}
\]

>>>>

\section{Stochastic Recurrence Equation}
One of the most important dynamical mechanisms that lead to regularly
varying r.v. is stochastic recursion of the following form:
\begin{equation}
  \label{eq:rhjyu}
  X_t = A_t X_{t-1} + B_t
\end{equation}
where $X_t$ is a $d$-dimensional random vector, $A_t$ is a $d\times d$
random matrix and $B_t$ is a $d$-dimensional vector, random or
deterministic.

@@

\section{GARCH models}
Introduced by Bollerslev in 1986, GARCH models have been hugely
popular in modelling volatility of financial time series and have
inspired numerous variants. Its original form is the following the
recurrence equation:
\[
\sigma_t^2 = \omega + \sum_{i=1}^p \alpha_i R_{t-i}^2 +
\sum_{j=1}^q \beta_j \sigma_{t-j}^2
\]
where $\R_t$ is a return series, e.g. stock returns, foreign exchange
rates, interest rates, etc; $\sigma_t^2$ is the variance of the
distribution of $R_t$ conditional on $\{(R_i,
\sigma_i^2)\}_{i=0}^{t-1}$. $\omega, \{\alpha_i\}_{i=1}^p,
\{\beta_i\}_{i=1}^q$ are constant parameters of the model. Clearly,
the GARCH($p,q$) recurrence equation is of the form of
\eqref{eq:rhjyu}. So, with appropriate conditions including
$\sum_{i=1}^p \alpha_i + \sum_{j=1}^q \beta_j < 1$, $\sigma_t^2$ is
shown to be a positive Harris Markov chain (cf. Bollerslev
\cite{bollerslev:1986} and Buraczewski et al
\cite{buraczewski:damek:mikosch:2016}), whose stationary distribution
has regularly varying tails. The tail index, call it $\xi$, is given by
\[
\lim_{n \to \infty} {1 \over n}\log\E\|A_n \cdots A_1\|^\xi = 0
\]
where $A_1, A_2, \dots, A_n$ are iid matrices whose entries are
functions of $\{\alpha_i\}_{i=1}^p$ and $\{\beta_i\}_{i=1}^q$:
\[
A_i =
\begin{pmatrix}
  \alpha_1 Z_{t-1}^2 + \beta_1 & \beta_2 & \cdots &
  \beta_{q-1} & \beta_q & \alpha_2 & \alpha_3 &
  \cdots & \alpha_{p-1} & \alpha_p\\
  1 & 0 & \cdots & 
  0 & 0 & 0 & 0 & \cdots & 0 & 0 \\
  \vdots & \vdots & \ddots & 
  \vdots & \vdots & \vdots & \vdots &
  \ddots & \vdots & \vdots \\
  0 & 0 & \cdots &
  0 & 0 & 0 & 0 & \cdots & 0 & 0 \\
  0 & 0 & \cdots &
  1 & 0 & 0 & 0 & \cdots & 0 & 0 \\
  Z_{t-1}^2 & 0 & \cdots &
  0 & 0 & 0 & 0 & \cdots & 0 & 0 \\
  0 & 0 & \cdots &
  0 & 0 & 1 & 0 & \cdots & 0 & 0 \\
  \vdots & \vdots & \ddots &
  \vdots & \vdots & \vdots & \vdots &
  \ddots & \vdots & \vdots \\
  0 & 0 & \cdots &
  0 & 0 & 0 & 0 & \cdots & 0 & 0 \\    
  0 & 0 & \cdots &
  0 & 0 & 0 & 0 & \cdots & 1 & 0 \\    
\end{pmatrix}
\]
